---
title: "stm"
format: docx
editor: visual
---

```{r}
# Installing the required libraries
library(ggplot2)
library(dplyr)
library(car)

```
# Women's National Volleyball Team Data Set
```{r}
kadin_oyuncular <- data.frame(
  Isim = c("Ece Karayilan", "Aylin Sarioglu", "Melis Gurkaynak", "Sena Gulbahar", "Zehra Gunes", "Hande Baladin"),
  Pozisyon = c("Pasor", "Smacor", "Orta Oyuncu", "Libero", "Orta Oyuncu", "Smacor"),
  Yas = c(24, 22, 26, 23, 25, 24),
  Boy = c(180, 185, 188, 172, 190, 182),
  MacBasinaOrtalamaSayi = c(5, 12, 8, 0, 9, 11)
)

```
# Function to display the first few rows of the data set
```{r}
head(kadin_oyuncular)
```
# EDA for female actors
```{r}
summary(kadin_oyuncular)# Function to display summary of data set
```
# EDA: Data Visualization with Graphs
# Age Distribution
```{r}
ggplot(kadin_oyuncular, aes(x=Yas)) + geom_histogram(binwidth=1, fill="skyblue", color="black") + 
  ggtitle("Age Distribution of Players") + xlab("Yaş") + ylab("Frekans")
```

```{r}
#Height Distribution by Position
ggplot(kadin_oyuncular, aes(x=Pozisyon, y=Boy)) + geom_boxplot(fill="lightgreen") + 
  ggtitle("Height Distribution by Position") + xlab("Position") + ylab("Size (cm)")
```

```{r}
# Average Numbers Per Match
plot(kadin_oyuncular$Yas, kadin_oyuncular$MacBasinaOrtalamaSayi, main="Average Points Per Match by Law", xlab="Yas", ylab="Average Numbers Per Match", pch=19)

```

```{r}
# # Regression Analysis: Relationship Between Age and Average Points Per Match
lm_model_kadin <- lm(MacBasinaOrtalamaSayi ~ Yas, data=kadin_oyuncular)
summary(lm_model_kadin)
```

```{r}
# ANOVA
anova_model_kadin <- aov(MacBasinaOrtalamaSayi ~ Pozisyon, data=kadin_oyuncular)
summary(anova_model_kadin) # ANOVA sonuçlarını gösterir
```

```{r}
# ANCOVA 
ancova_model_kadin <- aov(MacBasinaOrtalamaSayi ~ Pozisyon + Boy, data=kadin_oyuncular)
summary(ancova_model_kadin) # ANCOVA sonuçlarını gösterir
```

```{r}
ggplot(kadin_oyuncular, aes(x=Pozisyon, y=MacBasinaOrtalamaSayi, fill=Pozisyon)) + 
  geom_boxplot() + 
  ggtitle("Average Points Per Game by Position") + 
  xlab("Pozisyon") + ylab("Average Points Per Match") +
  theme(legend.title = element_blank())

```

```{r}
# ANCOVA 
kadin_oyuncular$TahminEdilenSayi <- predict(ancova_model_kadin)
ggplot(kadin_oyuncular, aes(x=Boy, y=MacBasinaOrtalamaSayi, color=Pozisyon)) + 
  geom_point() + 
  geom_line(aes(y=TahminEdilenSayi), linetype="dashed") + 
  ggtitle("Average Points Per Match by Position and Size") + 
  xlab("Boy") + ylab("Average Points Per Match (Actual and Predicted)") +
  theme(legend.title = element_blank())

```

```{r}
# Men's National Volleyball Team Data Set
erkek_oyuncular <- data.frame(
  Isim = c("Burak Gungor", "Ahmet Tumer", "Kerem Elice", "Deniz Kuzey", "Murat Yenipazar", "Serhat Coskun"),
  Pozisyon = c("Pasor", "Smacor", "Orta Oyuncu", "Libero", "Orta Oyuncu", "Smacor"),
  Yas = c(26, 24, 28, 22, 27, 25),
  Boy = c(192, 198, 195, 174, 200, 188),
  MacBasinaOrtalamaSayi = c(4, 13, 7, 1, 10, 12)
)
```

```{r}
# Display the first few rows of the data set
head(erkek_oyuncular)
```

```{r}
# EDA for male actors
summary(erkek_oyuncular)
```

```{r}
# Average Points Per Match by Age
plot(erkek_oyuncular$Yas, erkek_oyuncular$MacBasinaOrtalamaSayi, main="Average Points Per Match by Age", xlab="Age", ylab="Average Points Per Match", pch=19)

```

```{r}
# EDA: Data Visualization with Graphs

# Age Distribution

ggplot(erkek_oyuncular, aes(x=Yas)) + geom_histogram(binwidth=1, fill="skyblue", color="black") +
  ggtitle("Age Distribution of Players") + xlab("Age") + ylab("Frequency")
```

```{r}
# Regression Analysis: Relationship Between Age and Average Points Per Match
lm_model_erkek <- lm(MacBasinaOrtalamaSayi ~ Yas, data=erkek_oyuncular)
summary(lm_model_erkek)
```

```{r}
# ANOVA Analysis: Differences in Average Points Per Match by Position
anova_model_erkek <- aov(MacBasinaOrtalamaSayi ~ Pozisyon, data=erkek_oyuncular)
summary(anova_model_erkek)

```

#ANOVA  Visualization

```{r}
# Average Points Per Game by Position - Box Plot
ggplot(erkek_oyuncular, aes(x=Pozisyon, y=MacBasinaOrtalamaSayi, fill=Pozisyon)) +
  geom_boxplot() +
  ggtitle("Average Points Per Game by Position") +
  xlab("Pozisyon") + ylab("Average Points Per Match") +
  theme(legend.title = element_blank())

```

```{r}
#ANCOVA Analysis
# ANCOVA Analysis: Neck Position and Its Effect on Average Points Per Match
ancova_model_erkek <- aov(MacBasinaOrtalamaSayi ~ Pozisyon + Boy, data=erkek_oyuncular)
summary(ancova_model_erkek)

```

###ANCOVA visualization
```{r}
erkek_oyuncular$TahminEdilenSayi <- predict(ancova_model_erkek)

# Average Points Per Match by Position and Size - Scatter Plot
ggplot(erkek_oyuncular, aes(x=Boy, y=MacBasinaOrtalamaSayi, color=Pozisyon)) +
  geom_point() +
  geom_line(aes(y=TahminEdilenSayi), linetype="dashed") +
  ggtitle("Average Points Per Match by Position and Size") +
  xlab("Size") +
  ylab("Average Points Per Match (Actual and Predicted))")


```

### Combining Datasets and Adding the Gender Factor

```{r}
# Merge data sets for male and female players
kadin_oyuncular$Cinsiyet <- 'Kadın'
erkek_oyuncular$Cinsiyet <- 'Erkek'

# Merge by adding the gender variable
oyuncular <- rbind(kadin_oyuncular, erkek_oyuncular)
oyuncular$Cinsiyet <- as.factor(oyuncular$Cinsiyet)

head(oyuncular)


```

```{r}
lm_model <- lm(MacBasinaOrtalamaSayi ~ Cinsiyet, data=oyuncular)
summary(lm_model)

```

#### 2.ANOVA Analysis: Effects of Gender and Position on Average Points Per Match
```{r}
anova_model <- aov(MacBasinaOrtalamaSayi ~ Cinsiyet + Pozisyon, data=oyuncular)
summary(anova_model)

```

#### 3. ANCOVA Analysis: Effect of Gender and Position When Height is Controlled

```{r}
ancova_model <- aov(MacBasinaOrtalamaSayi ~ Cinsiyet + Pozisyon + Boy, data=oyuncular)
summary(ancova_model)

```

### 1.	Average Points Per Match by Gender: Box Plot

```{r}
ggplot(oyuncular, aes(x=Cinsiyet, y=MacBasinaOrtalamaSayi, fill=Cinsiyet)) +
  geom_boxplot() +
  ggtitle("Average Scores per Match by Gender") +
  xlab("Gender") +
  ylab("Average Scores per Match") +
  theme_minimal()


```

### 2. Interaction of Position and Gender: Box Plot

```{r}
ggplot(oyuncular, aes(x=Pozisyon, y=MacBasinaOrtalamaSayi, fill=Cinsiyet)) +
  geom_boxplot() +
  ggtitle("Average Scores per Match by Position and Gender") +
  xlab("Position") +
  ylab("Average Scores per Match") +
  theme_minimal() +
  theme(legend.title = element_blank())


```

### 3. Relationship Between Height and Average Points Per Match: Scatter Plot

```{r}
ggplot(oyuncular, aes(x=Boy, y=MacBasinaOrtalamaSayi, color=Cinsiyet)) +
  geom_point() +
  geom_smooth(method="lm", aes(group=Cinsiyet), se=FALSE, linetype="dashed") +
  ggtitle("Relationship Between Height and Average Scores per Match") +
  xlab("Height") +
  ylab("Average Scores per Match") +
  theme_minimal() +
  theme(legend.title = element_blank())


```
